A Container-Based Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-10-17 draft.tex
More proof reading

Multicontainer applications are deployed to ECS as a series of tasks {\em tasks}, and a single cluster can run multiple tasks in
parallel. Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can also specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt cpu\_shares} indicating the CPU units
assigned, in this case, $2048$, and the network configuration; in this case, we have two private 
subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
need to be supported by the currently available cluster. For instance, we need to have all 
the CPU units required by all tasks because if they are not available, the task initiation will fail.

The complete task configuration is generated by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example, to deploy the controller, setup, and Redis 
containers, we used the file in Listing ~\ref{code:compose}.

@@

Because all these instances run in a private subnet, we also run a
Bastion instance, a specially protected network node, in a
public subnet, from where we send the commands to execute the
experiments and later collect the results. We connected to this
instance through an {\texttt ssh} session from our workstation.
This is an {\texttt R5AD Large} instance, selected with
75GiB of storage, needed for keeping the logs, and more memory (16GiB)
for processing the logs in memory. The details of these EC2 instances and
their cost at the time of writing are shown in Table \ref{tab:ec2}.

@@

In the previous section, we have described an option for an EvoSwarm
deployment that can be used by the public to create reproductive
research in the cloud; in this section, we are using the same
framework to learn if the proposed solution can efficiently improve
population-based optimization algorithms' scalability and performance.
Hence, in the following sections, we provide answers to
the following questions:

\begin{itemize}
\item Can we improve the algorithm's execution time by adding populations
 and serverless functions (workers)  to the system and, in particular,
 what is its effect on the scalability of the system? % Improve with respect to what? Single population? 
                         %Please explain carefully the issues and how scaling is related to the number of populations. - JJ
                         % The question could be that, How scaling is related to number of populations and workers in the system ? - M

@@

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective benchmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark of the testbed utilizes 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.

The COCO framework offers several tools to compare the performance of
algorithms, generating data sets, tables, and reports for an experiment. There
is a repository\footnote{\url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}} 
of more than 200 results for the noiseless BBOB testbed, collected from 
BBOB workshops and special sessions between the years 2009 and 2019. The
EvoSwarm application includes an adapted version of the noiseless BBOB testbed,
compatible with the scripts of the framework, to compare with other algorithms
in the repository.

To test the heterogeneous multi-population capabilities,  we compare the
performance between a homogeneous and an ensemble of multi-populations, using
Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). For the GA
implementation, we used the DEAP library \cite{fortin2012deap} and for the PSO,
the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
open-source. In EvoSwarm these two algorithms are implemented as stateless
functions.

@@

Next, we show the parameters for each algorithm, with Table
\ref{tab:GAparams} for the GA and Table \ref{tab:PSOparams} for the
PSO. We obtained these parameters following the same method as in a
previous work \cite{garcia2017benchmarking}.  To obtain the
parameters, we tested first on the Rastrigin separable function with
five dimensions. After about fifteen experiments, the most challenging
targets were achieved for this particular function. We tested again
with functions one to three, and after obtaining favorable results,
the PSO and GA algorithm parameters were set. In the GA, we randomly
set the mutation and crossover probabilities to have more
heterogeneous workers; for these parameters, the range of values is
specified. We did not change these parameters during the experiments,
and only the number of populations, population size, and number of
generations were provided as parameters. This strategy has been
applied successfully for increasing the performance of
multi-population algorithms without requiring the setting of initial
parameters \cite{garcia2014randomized}.

The experiments can be easily reproduced by using the configuration
files the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:evoswarm:config}. 
As a first step, we can deploy the docker application using a
{\tt docker-compose.yml} file, as described in Section~\ref{sec:evoswarm:config} for a local
test, or by following the steps for a cloud setting as described in Section~\ref{cloud-aws}. 
Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that every population will run
the GA algorithm.

@@

Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of FEs we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops, that is, complete message processings, will be performed.
All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

We have deployed the EvoSwarm application in Amazon Elastic Container Service as 
described in Section~\ref{cloud-aws}, each EC2 instance was an Amazon Linux 2 AMI with 
ECS Agent 1.42.0 and Docker version 19.03.6-ce. Additionally the BBOB experiments 
were performed with COCO \cite{hansen2016coco} version bbob.v15.03
in Python, the plots were produced with version 2.3.2.

@@

As we mentioned before in Section~\ref{method}, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},
have used a similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
by simulating the work of nodes by using a sleep function and messages of different sizes.
In a population-based algorithm like EvoSwarm, several parameters can increase or 
decrease the workload of workers:

@@

From this description, it can be deduced that we give more importance in this
paper to the number of populations since they are the basic unit of
work, the message, that
can be executed in parallel. Together with the size of each population and the
number of generations, this parameter determines the maximum number of function
evaluations for a single iteration. We do not have control over each algorithm's
complexity, but future analyses could focus on balancing populations and number
of worker nodes according to this.

In this experiment, we evaluate how the number of populations and the
number of workers in the system are related to the scalability of the
system.  In the next section, we describe the experimental setup and a
discussion of the results follows, with respect to the speed-up in
terms of evaluations per second (Section~\ref{sec:speedup}), and
execution time (Section~\ref{sec:exec-time}).

@@

To answer the first research question, we need to evaluate the effect of the
number of populations and the number of workers on the scalability of the
system, and in order to achieve this we propose an experiment for
which we have selected $f_4$ (Skew
Rastrigin-Bueche separable) from the BBOB testbed, over 10, and 20 dimensions. This function has been used
because it is computationally demanding, and in higher dimensions has been
difficult for PSO \cite{el2009black} and GAs \cite{nicolau2009application} to
solve with the required FEs. As we are only comparing in terms of scalability,
the results from a single function can be better understood, since there are
fewer factors involved. Following the procedure described in
Section~\ref{setup}, we executed two sets of experiments with five, ten, and twenty
initial populations, repeating each experiment 30 times using 1, 2, 4, 8, and 16
workers. We kept the same maximum number of FEs, changing the relevant
parameters for this. See Table \ref{tab:params:10} for the complete list.

@@

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{eval_ratio}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$,
          with different number of workers and  populations. The $y$ axis is in log scale.}
  \label{fig:spworker:evalspersecond}
\end{figure}

When comparing the results in this section, we performed the Wilcoxon Signed-Rank test, with a
significance level of 5\%. We indicate with an asterisk those results that are
not significant throughout the tables of this section.  % Where? - JJ
% Answered - Mario
To measure the speed-up obtained by the addition of workers, we will use the
number of evaluations per second, instead of the time required to finish an
instance of a problem. The reason for this is that in some cases reaching the
desired target can require less work because we are using a stochastic search. 
We are going to measure the total increase of work done in a second by adding more workers.

>>>>

Multicontainer applications are deployed to ECS as a series of tasks {\em tasks},    and a single cluster can run multiple tasks in
parallel. Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can also specify a task's configuration 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt cpu\_shares} indicating the CPU units
assigned, in this case, $2048$, and the network configuration; in this case, we have two private 
subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
need to be supported by the currently available cluster. For instance, we need to have all 
the CPU units required by all tasks because if they are not available, the task initiation will fail.

The complete task configuration is generated on-the-fly by the {\texttt ecs compose command} 
that uses a {\texttt docker-compose.yml} file to generate 
the configuration. For example, to deploy the controller, setup, and Redis 
containers, we used the file in Listing ~\ref{code:compose}.

@@

Because all these instances run in a private subnet, we also run a
Bastion instance, a specially protected network node, in a
public subnet. From this instance, we send the commands to execute the
experiments and later collect the results. We connected to this
instance through an {\texttt ssh} session from our workstation.
This is an {\texttt R5AD Large} instance, selected with
75GiB of storage, needed for keeping the logs, and more memory (16GiB)
for processing the logs in memory. The details of these EC2 instances and
their cost at the time of writing are shown in Table \ref{tab:ec2}.

@@

In the previous section, we have described an option for an EvoSwarm
deployment that can be used by the public to create reproductive
research in the cloud. In this section, we are using the same
framework to learn if the proposed solution can efficiently improve
population-based optimization algorithms' scalability and performance.
Hence, in the following sections, we provide answers to
the following questions:

\begin{itemize}
\item Can we improve the algorithm's execution time by adding populations
 and serverless functions (workers)  to the system and, in particular,
 what is its effect on the system's scalability? % Improve with respect to what? Single population? 
                         %Please explain carefully the issues and how scaling is related to the number of populations. - JJ
                         % The question could be that, How scaling is related to number of populations and workers in the system ? - M

@@

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective benchmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark runs 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.

The COCO framework offers several tools to compare algorithms' performance,
generating data sets, tables, and reports for an experiment. There
is a repository\footnote{\url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}} 
of more than 200 results for the noiseless BBOB testbed, collected from 
BBOB workshops and special sessions between the years 2009 and 2019. The
EvoSwarm application includes an adapted version of the noiseless BBOB testbed,
compatible with the scripts of the framework, to compare with other algorithms
stored in the repository.

To test the heterogeneous multi-population capabilities,  we compare the
performance between a homogeneous and an ensemble of multi-populations, using
Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). 
We used the DEAP library \cite{fortin2012deap}, for the GA implementation. 
For the PSO algorithm, the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
open-source. In EvoSwarm, we implemented these two algorithms as stateless
functions.

@@

Next, we show each algorithm's parameters, with Table
\ref{tab:GAparams} for the GA and Table \ref{tab:PSOparams} for the
PSO. We obtained these parameters following the same method as in 
previous work \cite{garcia2017benchmarking}.  To obtain the
parameters, we tested first on the Rastrigin separable function with
five dimensions.  After about fifteen experiments, the algorithm 
achieved the most challenging targets for this particular function. We tested again
with functions one to three, and after obtaining favorable results, 
we set the PSO and GA parameters to these values. In the GA, we randomly
set the mutation and crossover probabilities to have more
heterogeneous workers; for these parameters, we specify the range 
of values in Table \ref{tab:GAparams}. We did not change these 
parameters during the experiments and only modified the number of
populations, population size, and generations. This strategy has been
applied successfully for increasing the performance of
multi-population algorithms without requiring the setting of initial
parameters \cite{garcia2014randomized}.

The experiments can be easily reproduced by using the configuration
files described in Section~\ref{sec:evoswarm:config}. 
As a first step, we can deploy the docker application using a
{\tt docker-compose.yml} file, as described in Section~\ref{sec:evoswarm:config} for a local
test, or by following the steps for a cloud setting as described in Section~\ref{cloud-aws}. 
Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$, there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that every population will run
the GA algorithm.

@@

Next, we specify a list indicating which of the 24 benchmark
functions we will test. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function we will teste. {\em Instances}  have a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of FEs we mentioned
earlier, for each dimension, we define the number of populations that will be
generated in the setup, and for each population, the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops, that is, the maximum number of messages taken from the queue.
All these parameters give us the maximum
number of FEs we will have as budget. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

We have deployed the EvoSwarm application in Amazon Elastic Container Service as 
described in Section~\ref{cloud-aws}, each EC2 instance was an Amazon Linux 2 AMI with 
ECS Agent 1.42.0 and Docker version 19.03.6-ce. Additionally, the BBOB experiments 
were performed with COCO \cite{hansen2016coco} version bbob.v15.03
in Python, and the plots were produced with version 2.3.2.

@@

As we mentioned before, in Section~\ref{method}, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},
have used a similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time, to scale on multiple
nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
by simulating the nodes' work by using a sleep function and messages of different sizes.
In a population-based algorithm like EvoSwarm, several parameters can increase or 
decrease the workload of workers:

@@

From this description, we can deduce that we give more importance 
to the number of populations since they are the basic unit of
work, the message, that can be executed in parallel. Together with each population's size and the
number of generations, this parameter determines the maximum number of function
evaluations for a single iteration. According to this, we do not have control 
over each algorithm's complexity, but future analyses could focus on balancing 
populations and the number of worker nodes.

In this experiment, we evaluate how the number of populations and workers 
are related to the system's scalability.
In the next section, we describe the experimental setup.A
discussion of the results follows, concerning the speed-up in
terms of evaluations per second (Section~\ref{sec:speedup}) and
execution time (Section~\ref{sec:exec-time}).

@@

To answer the first research question, we need to evaluate the effect of the
number of populations and the number of workers on the system's scalability.
To achieve this, we propose an experiment for
which we have selected $f_4$ (Skew Rastrigin-Bueche separable) from the BBOB testbed,
over 10 and 20 dimensions. This function has been used
because it is computationally demanding, and in higher dimensions, it has been
difficult for PSO \cite{el2009black} and GAs \cite{nicolau2009application} to
solve with the required number of FEs. As we are only comparing in terms of scalability,
the results from a single function can be better understood since fewer factors are involved.
Following the procedure described in Section~\ref{setup}, we executed two sets of 
experiments with five, ten, and twenty initial populations, repeating each 
experiment 30 times using 1, 2, 4, 8, and 16
workers. We kept the same maximum number of FEs, changing the relevant
parameters for this. See Table \ref{tab:params:10} for the complete list.

@@

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{eval_ratio}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$,
          with a different number of workers and populations. The $y$-axis is in log scale.}
  \label{fig:spworker:evalspersecond}
\end{figure}

When comparing the results in this section, we performed the Wilcoxon Signed-Rank test, with a
significance level of 5\%. We indicate with an asterisk those results that are
not significant throughout the tables of this section.  % Where? - JJ
% Answered - Mario
To measure the speed-up obtained by the addition of workers, we will use the
number of evaluations per second instead of the time required to finish an
instance of a problem. This is because in some cases reaching the
desired target can require less work because we are using a stochastic search. 
We are going to measure the total increase of work done in a second by adding more workers.
